### 2024-08-23 cognitive-bias.qmd
20240822 - affect heuristic

As described [above](#sec-cognitiveBiasOverview), heuristics are mental shortcuts that people use to handle the overwhelming amount of information to process.
Important heuristics used in judgment and prediction in the face of uncertainty include [@Tversky1974; @Critcher2014]:

- [availability heuristic](#sec-heuristicsAvailability)
- [representativeness heuristic](#sec-heuristicsRepresentativeness)
- [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment)
- [performance heuristic](#sec-heuristicsPerformance)

@@

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to the person's existing mental prototypes.
For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

>>>>

As described [above](#sec-cognitiveBiasOverview), heuristics are mental shortcuts that people use to handle the overwhelming amount of information to process.
Important heuristics used in judgment and prediction in the face of uncertainty include [@Kahneman2011; @Tversky1974; @Critcher2014]:

- [availability heuristic](#sec-heuristicsAvailability)
- [representativeness heuristic](#sec-heuristicsRepresentativeness)
- [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment)
- [affect heuristic](#sec-heuristicsAffect)
- [performance heuristic](#sec-heuristicsPerformance)

@@

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

### Affect Heuristic {#sec-heuristicsAffect}

The *affect heuristic* refers to the tendency for a person to make judgments or predictions based on their emotions and feelings—such as whether they like or dislike something—with little deliberation or reasoning [@Kahneman2011].
For instance, if a manager's favorite team is the Dallas Cowboys, they may hold a distaste for players on teams that are rivals of the Cowboys, such as the Philadelphia Eagles.
The affect heuristic may lead them to avoid drafting Eagles players even though some Eagles players may perform well.
